---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(cart)
library(terra)
library(geojsonsf)
library(jsonlite)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
```

To parameterise a malariasimulation run for a specific context, we often need
to gather and summarise spatial data, provided in the form of rasters. We can
use cart to help get commonly used versions of these data. This can be

1. Population estimates from [WorldPop](https://www.worldpop.org/).
2. Vector species data from [Malaria Atlas Project](https://malariaatlas.org/).
3. Prevalence maps from [Malaria Atlas Project](https://malariaatlas.org/).
4. Spatial limits of transmission maps from [Malaria Atlas Project](https://malariaatlas.org/).


Let's have a look at an example country. We can get the admin 2 unit boundaries
for Togo from [geoBoundaries](https://www.geoboundaries.org/index.html).

```{r, warning=FALSE}
# Get Togo admin 1 polygons from geoboundaries API
shape_file <- jsonlite::fromJSON("https://www.geoboundaries.org/api/current/gbOpen/TGO/ADM2/")
# As simple feature
sf_poly <- geojsonsf::geojson_sf(shape_file$gbAuthoritative$simplifiedGeometryGeoJSON)
# As spatVector
spatvec_poly <- as(sf_poly, "SpatVector")
```

Now we can get the raster information we need for Togo

```{r, warning=FALSE, message=FALSE, fig.width=7, fig.height=4}
# Pull data
data <- pull_cart(iso3c = "TGO", year = 2019)
data

# Visualise some outputs
par(mfrow = c(1, 3))
terra::plot(data, y = 1, col = viridis::magma(30), main = "Population")
lines(spatvec_poly, col = "white")
terra::plot(data, y = 2, col = viridis::plasma(30), main = "PfPr", range = c(0, 1))
lines(spatvec_poly, col = "white")
terra::plot(data, y = 4, main = "Funestus", range = c(0, 1))
lines(spatvec_poly, col = "white")
par(mfrow = c(1, 1))
```

Let's extract those data
```{r, warning=FALSE}
extracted_data <- unpack_cart(sf_poly, data)
```
There are lots of post-processing steps we might want to do. For example, let's
estimate the population weighted falciparum prevalence for each administrative unit in 
Togo

```{r, fig.width=6, fig.height=6}
summary_data <- extracted_data %>%
  dplyr::select(shapeName, pop, pfpr) %>%
  tidyr::unnest(cols = c(pop, pfpr)) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(shapeName) %>%
  dplyr::summarise(pfpr = round(weighted.mean(pfpr, pop), 3))

fc <- dplyr::left_join(sf_poly, summary_data, by = "shapeName") %>%
  dplyr::select(pfpr)

ggplot(fc, aes(fill = pfpr)) + 
  geom_sf() +
  scale_fill_viridis_c(option = "plasma", name = "", limits = c(0, 0.42)) +
  theme_bw() +
  ggtitle("Aggregated PfPr")
```

Other common operations we might want to perform would may include:

1. Estimating population at risk, by setting the population in pixels outside
of the spatial limits of transmission to equal 0
2. Summarising the relative abundance of the top N vector species 
