---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(cart)
library(terra)
library(geojsonsf)
library(jsonlite)
library(dplyr)
library(tidyr)
```

To parameterise a malariasimulation run for a specific context, we often need
to gather and summarise spatial data, provided in the form of rasters. We can
use cart to help get commonly used versions of these data. This can be

1. Population estimates from [WorldPop](https://www.worldpop.org/).
2. Vector species data from [Malaria Atlas Project](https://malariaatlas.org/).
3. Prevalence maps from [Malaria Atlas Project](https://malariaatlas.org/).
4. Spatial limits of transmission maps from [Malaria Atlas Project](https://malariaatlas.org/).

To start we can supply the iso3c code for Togo and a recenbt year.

```{r, message=FALSE, fig.width = 7, fig.height = 7}
data <- pull_cart(iso3c = "TGO", year = 2019)
terra::plot(data)
```

Next, we're going to extract the values for a given simple feature polyon set.
Usually this would be nationally, or for some sub-national regions. Here, we're
going to use the admin 1 units in Togo from [geoBoundaries](https://www.geoboundaries.org/index.html).

```{r, warning=FALSE}
shape_file <- jsonlite::fromJSON("https://www.geoboundaries.org/api/current/gbOpen/TGO/ADM1/")
fc <- geojsonsf::geojson_sf(shape_file$gbAuthoritative$simplifiedGeometryGeoJSON)
extracted_data <- unpack_cart(fc, data)
```
There are lots of post-processing steps we might want to do. For example, let's
estimate the population weighted falciparum prevalence for each administrative unit in 
Togo

```{r, fig.width = 7, fig.height = 7}
summary_data <- extracted_data %>%
  dplyr::select(shapeName, pop, pfpr) %>%
  tidyr::unnest(cols = c(pop, pfpr)) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(shapeName) %>%
  dplyr::summarise(pfpr = round(weighted.mean(pfpr, pop), 3))

fc <- dplyr::left_join(fc, summary_data, by = "shapeName") %>%
  dplyr::select(pfpr)
plot(fc)
```
